---
title: "Stat 222 - Exercise 1"
author: "kclazaro"
date: "2023-02-21"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

**Exercise.** Refer to the height-weight data set in the Lesson 2 Folder in the Class Drive. Assume the data is taken from certain population through random sampling. Let X,Y be the height and weight, respectively, of a randomly selected person in the population. Find the observed value of the following statistics in Excel or R.

1. $\overline{X}$
2. $S^2_Y$ (sample variance of Y)
3. $Y_{(2)}$
4. $M{'}_{X,2}$ (second sample raw moment of X)

Using the data set, we get the following summary statistics.
```{r}
#Load the 'readxl' library to be able to parse Excel files. Then we assign the variable name 'weight_height' to the Excel data set. Lastly, we run the summary statistics of the data set.
library(readxl)
weight_height <- read_excel("Downloads/weight-height.xlsx")
summary(weight_height)
```

To get the mean of the height column, we use the following code:
```{r}
# Get the mean of the Height column by column name
xbar <- mean(weight_height$Height)
```

This means that the mean height of the data set is...
```{r echo=FALSE}
xbar
```

For the sample variance of Y, we use the following code:
```{r}
# Get the variance of the Weight column by column name
var_y <- var(weight_height$Weight)
```

Using this, the variance of the weight column is...
```{r echo=FALSE}
var_y
```

In order to obtain the second lowest value in the Weight column, we will use the following code:
```{r}
# Sort the Weight column from lowest to highest, and we assign it to the 'weight_order' variable.
weight_order <- sort(weight_height$Weight)
```

From this, it's easy to obtain the second smallest value, which is...
```{r echo=FALSE}
weight_order[2]
```

Lastly, to get the second sample raw moment of height, we can use the moments package.
```{r}
# Load the moments package then get the second raw moment by specifying the order.
library(moments)
second_raw_moment_height <- moment(weight_height$Height, order=2)
```

Which will then give us the second sample raw moment for the height column...
```{r echo=FALSE}
second_raw_moment_height
```

In summary, the following are the observed values for the statistics:
```{r echo=FALSE}
library(kableExtra)

dt <- read_excel("Downloads/stat222-exercise1.xlsx")
kbl(dt) %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

